---
title: "Final Project"
author: "Bill Huang"
date: "4/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
cat("\014")
```

```{r}
data <- data.frame(read.csv("Life Expectancy Data.csv"))
Status2 <- ifelse(data$Status == "Developed", 1,0)
data <- cbind(data, Status2)
df <- na.exclude(data)
library(mgcv)
```

```{r}
plot(data$Alcohol, data$Life.expectancy)
fit <- lm(Life.expectancy ~ Alcohol, data = data)
abline(fit$coefficient[1], fit$coefficient[2], col="red")
fit_log <- lm(log(Life.expectancy) ~ Alcohol, data = data)

```


```{r}
fit_gam <- gam(Life.expectancy ~ s(Alcohol), data = df)
plot(df$Alcohol, df$Life.expectancy)
lines(df$Alcohol, fitted(fit_gam), col="red")
```

```{r}
df_2000 <- data[data$Year=="2000",]
fit_2000 <- lm(df_2000$Life.expectancy ~ df_2000$Alcohol)
plot(df_2000$Alcohol, df_2000$Life.expectancy, xlab="Alcohol consumption in 2000", ylab="Life expectancy in 2000")
abline(a = fit_2000$coefficient[1], 
       b = fit_2000$coefficient[2], 
       col="red")
summary(fit_2000)
```

```{r}
df_2001 <- data[data$Year=="2001",]
fit_2001 <- lm(df_2001$Life.expectancy ~ df_2001$Alcohol)
plot(df_2001$Alcohol, df_2001$Life.expectancy, xlab="Alcohol consumption in 2001", ylab="Life expectancy in 2001")
abline(a = fit_2001$coefficient[1], 
       b = fit_2001$coefficient[2], 
       col="red")
summary(fit_2001)
```

```{r}
df_2002 <- data[data$Year=="2002",]
fit_2002 <- lm(df_2002$Life.expectancy ~ df_2002$Alcohol)
plot(df_2002$Alcohol, df_2002$Life.expectancy, xlab="alcohol consumption in 2002", ylab="Life expectancy in 2002")
abline(a = fit_2002$coefficient[1], 
       b = fit_2002$coefficient[2], 
       col="red")
summary(fit_2002)
```


```{r}
df_2012 <- data[data$Year=="2012",]
fit_2012 <- lm(df_2012$Life.expectancy ~ df_2012$Alcohol)
plot(df_2012$Alcohol, df_2012$Life.expectancy, xlab="alcohol consumption in 2012", ylab="Life expectancy in 2012")
abline(a = fit_2012$coefficient[1], 
       b = fit_2012$coefficient[2], 
       col="red")
summary(fit_2012)
```


```{r}
df_2013 <- data[data$Year=="2013",]
fit_2013 <- lm(df_2013$Life.expectancy ~ df_2013$Alcohol)
plot(df_2013$Alcohol, df_2013$Life.expectancy, xlab="alcohol consumption in 2013", ylab="Life expectancy in 2013")
abline(a = fit_2013$coefficient[1], 
       b = fit_2013$coefficient[2], 
       col="red")
summary(fit_2013)
```


```{r}
df_developing <- data[data$Status=="Developing", ]
df_developed <- data[data$Status=="Developed", ]
fit_developing <- lm(df_developing$Life.expectancy ~ df_developing$Alcohol)
plot(df_developing$Alcohol, df_developing$Life.expectancy)
abline(a = fit_developing$coefficient[1],
       b = fit_developing$coefficient[2],
       col = "red")
summary(fit_developing)

fit_developed <- lm(df_developed$Life.expectancy ~ df_developed$Alcohol)
plot(df_developed$Alcohol, df_developed$Life.expectancy)
abline(a = fit_developed$coefficient[1],
       b = fit_developed$coefficient[2],
       col = "red")
summary(fit_developed)


```

```{r}
df_developing_2000 <- df_developing[df_developing$Year == "2000",]
plot(df_developing_2000$Alcohol, df_developing_2000$Life.expectancy)
fit_developing_2000 <- lm(df_developing_2000$Life.expectancy ~ df_developing_2000$Alcohol)
abline(a = fit_developing_2000$coefficients[1],
       b = fit_developing_2000$coefficients[2],
       col="red")
summary(fit_developing_2000)

df_developed_2000 <- df_developed[df_developed$Year == "2000",]
plot(df_developed_2000$Alcohol, df_developed_2000$Life.expectancy)
fit_developed_2000 <- lm(df_developed_2000$Life.expectancy ~ df_developed_2000$Alcohol)
abline(a = fit_developed_2000$coefficients[1],
       b = fit_developed_2000$coefficients[2],
       col="red")
summary(fit_developed_2000)


```

```{r}
df_developing_2005 <- df_developing[df_developing$Year == "2005",]
plot(df_developing_2005$Alcohol, df_developing_2005$Life.expectancy)
fit_developing_2005 <- lm(df_developing_2005$Life.expectancy ~ df_developing_2005$Alcohol)
abline(a = fit_developing_2005$coefficients[1],
       b = fit_developing_2005$coefficients[2],
       col="red")
summary(fit_developing_2005)


df_developed_2005 <- df_developed[df_developed$Year == "2005",]
plot(df_developed_2005$Alcohol, df_developed_2005$Life.expectancy)
fit_developed_2005 <- lm(df_developed_2005$Life.expectancy ~                                               df_developed_2005$Alcohol)
abline(a = fit_developed_2005$coefficients[1],
       b = fit_developed_2005$coefficients[2],
       col="red")
summary(fit_developed_2005)

```

```{r}
fit_gam <- gam(Life.expectancy ~ s(Alcohol), data = df_developing)
summary(fit_gam)
plot(fit_gam, shade=T)
```



```{r}

fit_null <- lm(Life.expectancy ~ 1, data = df)
fit_full <- lm (Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol +
                percentage.expenditure + Hepatitis.B + Measles + BMI + 
                under.five.deaths + Polio + Total.expenditure + Diphtheria +                      HIV.AIDS + GDP + Population + thinness..1.19.years +                              thinness.5.9.years  + Income.composition.of.resources + Schooling +                Status2, data = df)
step(fit_null , direction = "forward", scope = formula(fit_full))



```
     After computing forward method of AIC, it appears that we had excluded 
"Hepatitis.B", "Measles", "Polio", "GDP", "Population", "as well as 
"prevalence of thinness among children and adolescents between 10 to 19" from all 
the variables.
      The best model by forward method : Life.expectancy ~ Schooling + HIV.AIDS + 
Adult.Mortality + Income.composition.of.resources + 
percentage.expenditure + BMI + Diphtheria + under.five.deaths + 
infant.deaths + thinness.5.9.years + Total.expenditure + Alcohol
```{r}
anova(fit_null, fit_full)
```



```{r}
y <- df$Life.expectancy
X <- model.matrix(~Adult.Mortality + infant.deaths + Alcohol + 
                    percentage.expenditure + BMI + under.five.deaths + 
                    Total.expenditure + Diphtheria + HIV.AIDS + thinness.5.9.years                    + Income.composition.of.resources + Schooling + Status2, 
                  data = df)
fit_best <- lm(y ~ X)
summary(fit_best)
XtX = t(X)%*%X
invXtX = solve(XtX)
Xty = t(X)%*%y
beta_hat = invXtX%*%Xty
plot(fit_best)

```


```{r}
split <- sample(c(rep(0, 0.7*nrow(df)),
                  rep(1, 0.3*nrow(df))))
df_train <- df[split == 0,]
df_test <- df[split == 1, ]
fit_train <- lm(Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol + 
                    percentage.expenditure + BMI + under.five.deaths + 
                    Total.expenditure + Diphtheria + HIV.AIDS +                                       thinness.5.9.years + Income.composition.of.resources +                            Schooling + Status2, 
                    data = df_train)
predict <- predict(fit_train,
                   newdata = df_test,
                   interval = "predict",
                   level = 0.9)
mean(abs(df_test$Life.expectancy - predict))
max(df_test$Life.expectancy - predict)
mean((df_test$Life.expectancy - predict)^2)
hist(abs(df_test$Life.expectancy - predict))

```



```{r}
fit_gam2 <- gam(y ~ Adult.Mortality + infant.deaths + Alcohol + 
                    percentage.expenditure + BMI + under.five.deaths + 
                    s(Total.expenditure) + Diphtheria + HIV.AIDS +                                    s(thinness.5.9.years) + Income.composition.of.resources +                         Schooling, 
                    data = df)
plot(fit_gam2, shade=T)
```

```{r}
library(glmnet)
fit.cv1 <- cv.glmnet(X,
                     y,
                     alpha = 0)
coef(fit.cv1)
plot(fit.cv1)
fit.cv1$lambda.min


fit.cv2 <- cv.glmnet(X,
                     y,
                     alpha = 1)
plot(fit.cv2)
fit.cv2$lambda.min
coef(fit.cv2)
```

